---
title: "Cleaning strategies - Oct 11"
output: 
---

LIBRARIES:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(readxl)
```

DATA (located in Files on Canvas):

```{r}
# deaths file h/t Mary Jo Webster, data editor of the Minnesota Star Tribune
mn_deaths <- read_csv("data/MN_opiate_deaths.csv")

# disney file h/t Andrew Ba Tran, investigative data reporter at the Washington Post
disney <- read_csv("data/disney_movies_total_gross.csv")

# use read_excel() to import an Excel spreadsheet
salaries <- read_excel("data/Bloomington Salaries.xlsx", sheet=1)
```

TASKS: `disney`

Check categorical variables for cleanliness: 
```{r}
disney %>% count(genre)
disney %>% count(MPAA_rating)
```

1. convert the dollars column to numbers
2. convert the date column to a true date

```{r}
disney <- disney %>% 
  mutate(release_date = mdy(release_date))

# alternative to coercing these columns into numbers on import (which works)
disney <- disney %>% 
  mutate(total_gross = parse_number(total_gross),
         inflation_adjusted_gross = parse_number(inflation_adjusted_gross))
```

Which disney movie is the highest-grossing movie of all time? 
```{r}
disney %>% arrange(desc(inflation_adjusted_gross))
```


TASKS: `salaries`

How many NAs are in the dataset? 
Remember how to ignore NAs
```{r}
#How many NAs? 
summary(salaries)

#Ignore NAs in summary functions
salaries %>% summarise(total = sum(overtime_oncall, na.rm=T))
```

Check categorical variables
```{r}
salaries %>% count(department)
salaries %>% count(job_title)
```

1. Turn all "Utilities" sub-departments into one department

```{r}
salaries <- salaries %>% 
  mutate(new_department = ifelse(grepl("Utilities",department,ignore.case=T),"Utilities",department), .after = department)
```
 
Which department has the most employees? 
```{r}
salaries %>% 
  count(new_department) %>% 
  arrange(desc(n))
```

Which department has the highest average salary?
Which department has the highest average total compensation? 
```{r}
# using total compensation
salaries %>% 
  group_by(new_department) %>% 
  summarise(avg_total = mean(total_comp)) %>% 
  arrange(desc(avg_total))

# what about salary?
salaries %>% 
  group_by(new_department) %>% 
  summarise(avg_sal = mean(salary_2021, na.rm=T)) %>% 
  arrange(desc(avg_sal))

# check hourly pay
salaries %>% 
  group_by(new_department) %>% 
  summarise(avg_hourly = mean(hourly_rate, na.rm=T)) %>% 
  arrange(desc(avg_hourly))
```

Which department pays out the most overtime? 
```{r}
salaries %>% 
  group_by(new_department) %>% 
  summarise(tot_over = sum(overtime_oncall, na.rm=T)) %>% 
  arrange(desc(tot_over))
```

Who got the most overtime in that department? 
```{r}
salaries %>% 
  filter(new_department=="Police") %>% 
  mutate(overtime_pct = overtime_oncall/total_comp) %>% 
  arrange(desc(overtime_pct))
```


TASKS: `mn_deaths`

- check if there are duplicates (using StateID & using names/dob)
- check / clean up HISPANICETHNICITY
- convert BIRTDATE & DEATHDATE
- note: a subset of deaths related to opiates
- check RACE, GENDER, PLACETYPE

*Check for duplicates*
```{r}
mn_deaths %>% 
  count(STATEID) %>% 
  arrange(desc(n))
# no duplicates in STATEID

mn_deaths %>% 
  count(FIRSTNAME, MIDDLENAME, LASTNAME, BIRTHDATE) %>% 
  arrange(desc(n))
# no duplicates in names / DOB

mn_deaths %>% 
  count(LASTNAME, BIRTHDATE) %>% 
  arrange(desc(n))
```

*Check HISPANICETHNICITY*
```{r}
mn_deaths %>% 
  count(HISPANICETHNICITY)
# needs cleanup

mn_deaths %>% 
  count(RACE)
# unique groups not matching Census but not messy

mn_deaths %>% 
  count(GENDER)
# binary but clean

mn_deaths %>% 
  count(PLACETYPE)
# not standardized
```

*CLEANUP*
```{r}
# convert birthdate and deathdate


```

