---
title: "Dispatch"
output: 
---

Analysis of Columbia, MO police dispatch data from https://www.como.gov/CMS/911dispatch/police.php
Downloaded on 9/6/2023 for the previous 3 years (9/6/2020 - 9/6/2023)

```{r}
library(tidyverse)
```

Bring in the data: 
```{r}
?read_csv

dispatch <- read_csv("data/como_dispatch.csv")
```

Main questions and vetting the data. 
1. How many rows of data? *85,812*
2. What is one row? *one row is one incident where officer(s) were dispatched*
3. What's in each column? 

```{r}
# Check InNum: 
dispatch %>% 
  count(InNum) %>% 
  arrange(desc(n)) %>% 
  filter(n > 1)

dispatch %>% 
  filter(InNum == 2022208992)
# there's some dirty data here - duplicate rows with slight variations (maybe a correction creates a duplicate record?) - ask PD
```


Homework questions: 

Go through each column using the count() function: 
Is there any messy data? Are there any columns that need to be converted and/or cleaned for analysis? 
How many categories are in the ExtNatureDisplayName column? Which is the most common? 
Which day of the week has the most dispatches? 
Which PolArea had the most dispatches? 
Think like a journalist and come up with five questions you want to ask this data. 
