---
title: "WNBA analysis - aug 28"
output: 
---

Analysis of WNBA data from HerHoopStats.com and ESPN
Player salaries for 2023 and stats from 2022

```{r}
library(tidyverse)
```

Import the file wnba2023.csv from our data folder
```{r}
wnba <- read_csv("data/wnba2023.csv")
```

Questions of the data: 
1. How many rows of data are there? 148 rows
2. What is one row? 1 row is 1 wnba player (as of pre-season)
3. What is in every column
  salary is number which is good
  height needs to be converted to a number if we want to use it
  jersey number is stored as text
  
Run-through of the basic functions for select, filter, arrange:
```{r}
wnba %>% 
  select(player_name, salary_2023) %>% 
  filter(salary_2023 > 100000) %>% 
  arrange(desc(salary_2023))

# this will return the exact same results, even though the order is different:
wnba %>% 
  arrange(desc(salary_2023)) %>% 
  filter(salary_2023 > 100000) %>% 
  select(player_name, salary_2023)

# salaries in descending order by team: 
wnba %>% 
  arrange(team, desc(salary_2023))

```

Run through summary numbers: 
```{r}
# the $ is base R vernacular and identifies a column within a dataset to sum, etc
sum(wnba$salary_2023)
mean(wnba$salary_2023)
median(wnba$salary_2023)

# here's how we use these functions within the tidyverse scheme
wnba %>% summarise(sum(salary_2023))
wnba %>% summarise(mean(salary_2023))
wnba %>% summarise(median(salary_2023))

# string them all together
wnba %>% 
  summarise(sum(salary_2023), mean(salary_2023), median(salary_2023))
```


Run through group_by: 
```{r}

wnba %>% 
  group_by(team) %>% 
  summarise(sum_sal = sum(salary_2023), mean_sal = mean(salary_2023), median_sal = median(salary_2023)) %>% 
  arrange(desc(mean_sal))

```


Homework questions: 
1. Which position has the highest average (mean) salary? 
2. How many players play for each team? hint: the summary function you want to use is n()
3. Which college has sent the most players to the wnba? 
4. How many players are from outside of the U.S.?
5. Do international players make a higher or lower average salary than players from the US?
6. Which position is the most common? 